---
title: "p8105_hw3_js5431"
author: "J Shearston"
date: "October 6, 2018"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))

```


## Problem #1

#### Data Import and Cleaning

```{r load & clean BRFSS}

library(p8105.datasets)

brfss = p8105.datasets::brfss_smart2010 %>% 
  janitor::clean_names() %>%
  rename(state = locationabbr, 
         county = locationdesc) %>% 
  filter(topic == "Overall Health") %>% 
  arrange(year, state, county) %>% 
  mutate(as.factor(response),
            fct_relevel(response, 
                        c('Excellent', 'Very good', 'Good', 'Fair', 'Poor')))


```


#### Questions and Responses

* In 2002, which states were observed at 7 locations?
    + *Response*

* Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.
    + *Response*

* Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.
    + *Response*

* For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.
    * *Response*


## Problem #2